The Correctness of Algebraic File Synchronization

### 2015-11-30 2015/paper.tex
New notation for commands

\newcommand{\setv}{\mathcal{V}} % Set of values
\newcommand{\setvx}[1]{\mathcal{V}_{#1}}
\newcommand{\setf}{\setvx{F}} % Set of file values
\newcommand{\setd}{\setvx{D}} % Set of directory values
\newcommand{\setb}{\setvx{\empt}} % Set of empty values
\newcommand{\setn}{\mathcal{N}} % Set of nodes / paths
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\topnode}{\mathtt{None}} % return value of parent()
\newcommand{\fsbroken}{\mathtt{Broken}} % broken FS value
\newcommand{\FS}{\Phi} % a filesystem
\newcommand{\GS}{\Xi} % another filesystem
\newcommand{\nn}{n^*} % parent node

\newcommand{\cbrk}{\mathtt{break}} % 'break' command
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyx}{\fscommand{Y}{X}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\cxw}{\fscommand{X}{W}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}
\newcommand{\cqq}{\fscommand{Q}{Q}}

\newcommand{\typeset}{\{\fsregcommandchar{D},\fsregcommandchar{F},\empt\}} % Set of types in commands
\newcommand{\cc}{\circ} % Command / sequence concatenation
\newcommand{\descendant}{\prec}
\newcommand{\descendantEq}{\preceq}
\newcommand{\ancestor}{\succ}
\newcommand{\ancestorEq}{\succeq}

@@

In this paper, we write $F$ for an arbitrary element of $\setf$, and $D$ for an arbitrary element
of $\setd$. $V$ is a value from $\setv$ and $n$ and $m$ are nodes in $\setn$.

@@

We therefore propose to have a command for each pair of types.
For want of a better system, we will name our commands by concatenating
two of $\fsregcommandchar{D}$, $\fsregcommandchar{F}$ and $\fsregcommandchar{\empt}$, 
where the left sign denotes the type of value
required in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example, $mkdir(n,D)$ is $\cbd(n,D)$, and $rmdir(n)$ is $\cdb(n)$.

@@

We note that a command or a sequence of commands is applied to a filsystem
by prefixing the command or sequence to it, for example: $\cbrk\FS$, $\cbd(n,D)\FS$, 
or $S\FS$ if $S$ is a sequence of commands.

@@

The exact behavior of our commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
All other commands have the generic form $\cxy(n,V_Y)$ where
$X$ and $Y$ are types in $\typeset$,
$n\in\setn$ and $V_Y$ is a value of the appropriate type from $\setvx{Y}$.
At times we omit $V_Y$ if there is only one suitable value.
\[ \cxy(n,V_Y)\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[V/n] \mbox{~violates the tree property}\\
   \FS[V/n] &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

In general, when we describe multiple or unknown commands, we may substitute one or both
of their types with variables, as we did with $\cxy(n,V_Y)$. In this context,
if we write that $\cxy=\czw$, we mean that the input and output types
of the commands are the same ($X=Z$ and $Y=W$), while a full equiality 
($\cxy(n,V_Y)=\czw(n,V_W)$)
implies
that their output value is also the same.

@@

The commands we aim to investigate do not occur in isolation:
for example, when comparing filesystems to determine the commands needed to transform one into the other,
we receive a set of commands with uncertain ordering.
Applying commands to a filesystem, however, happens in time,
so for this we need a sequence of commands.
We use $\cc$ to concatenate commands to form a sequence, or concatenate sequences to form a longer sequence:
$S=\cbd(n)\cc \cbf(m,F)$, or $S'=S\cc\cbb(o)$.
We write $\emptyseq$ for the empty sequence.

@@

Pairs of commands in general have the form
\[ \cxy(n,V_Y)\cc  \czw(m,V_W) \]
where $X,Y,Z,W\in\typeset$, $n,m\in\setn$, 
and values are of the appropriate type: 
$V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$.

\begin{myax}\label{ax_separate_commute}
Commands on unrelated nodes commute:
$\cxy(n,V_Y)\cc \czw(m,V_W) \equiv \czw(m,V_W)\cc \cxy(n,V_Y)$ where $n\unrel m$.
\end{myax}

\begin{myax}\label{ax_separate_nobreaks}
Commands on unrelated nodes do not break every filesystem:
$\cxy(n,V_Y)\cc \czw(m,V_W) \nequiv \cbrk$ where $n\unrel m$.
\end{myax}

\begin{myax}\label{ax_same_breaks}
Commands on the same node break every filesystem if their types are incompatible:
$\cxy(n,V_Y)\cc \czw(n,V_W) \equiv \cbrk$ where $Y\ne Z$.
\end{myax}

\begin{myax}\label{ax_same_emptyseq}
Commands on the same node simplify:
$\cxy(n,V_Y)\cc \czw(n,V_W) \eqext \emptyseq$ where $Y=Z$ and $X=W=\fsregcommandchar{\empt}$ 
or $X=W=\fsregcommandchar{D}$.
\end{myax}

\begin{myax}\label{ax_same_singlec}
Commands on the same node simplify:
$\cxy(n,V_Y)\cc  \czw(n,V_W) \equiv \cxw(n,V_W)$ where $Y=Z$ and $XW\neq\cbb$ and $XW\neq\cdd$.
\end{myax}

\begin{myax}\label{ax_distantrel_breaks}
Commands on distant relatives break all filesystems:
$\cxy(\nn,V_Y)\cc  \czw(n,V_W) \equiv \cbrk$
and $\czw(n,V_W)\cc \cxy(\nn,V_Y) \equiv\cbrk$
where $\nn\descendant n$ and $\nn\neq\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$.
\end{myax}

\begin{mydef}{Construction pair.}
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parent(n)$ 
and the commands are one
of the following:
   \begin{gather*}
   % bd>bf
            \cbd(\nn, D)\cc  \cbf(n, F) \\
   % bd>bd
            \cbd(\nn, D_1)\cc  \cbd(n, D_2) \\
   % fd>bf
            \cfd(\nn, D)\cc  \cbf(n, F) \\
   % fd>bd
            \cfd(\nn, D_1)\cc  \cbd(n, D_2)
   \end{gather*}
\end{mydef}

\begin{myax}\label{ax_directchild_breaks}
All other commands on a child break every filesystem:
$\cxy(\nn,V_Y)\cc  \czw(n,V_W) \equiv \cbrk$ where $\nn=\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.
\end{myax}

\begin{mydef}{Destruction pair.}
A pair of commands on nodes $n$ and $\nn$ is a destruction pair if $\parent(n)=\nn$ and the commands are one
of the following:
   \begin{gather*}
   % fb<db ??
            \cfb(n)\cc  \cdb(\nn) \\
   % fb<df ??
            \cfb(n)\cc  \cdf(\nn, F) \\
   % db<db ??
            \cdb(n)\cc  \cdb(\nn) \\
   % db<df ??
            \cdb(n)\cc  \cdf(\nn, F)
   \end{gather*}
\end{mydef}

\begin{myax}\label{ax_directparent_breaks}
All other commands on a parent break every filsystem:
$\cxy(n,V_Y)\cc  \czw(\nn,V_W) \equiv \cbrk$ where $\parent(n)=\nn$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.
\end{myax}

\begin{mydef}{Assertion command.}
A command is an assertion command if
for every filesystem it either breaks it or leaves it in the same state.
In other words, assertion commands have the same input and output type,
and there must also be only one possible value for their output.
Accordingly, in our current algebra, $\cbb(n)$ and $\cdd(n)$ are the only two assertion commands,
and we do not note their output value as it is determined.
\end{mydef}

\begin{myax}\label{ax_child_assert}
An assertion command can be added on a descendant node:
$\cbb(n)\cc \cxy(\nn,V_Y) \equiv \cxy(\nn,V_Y) \equiv \cxy(\nn,V_Y)\cc \cbb(n)$ 
where $\nn\descendant n$ and $\cxy\neq\cdd$.
\end{myax}

\begin{myax}\label{ax_parent_assert}
An assertion command can be added on an ancestor node:
$\cdd(\nn)\cc \cxy(n,V_Y) \equiv \cxy(n,V_Y) \equiv \cxy(n,V_Y)\cc \cdd(\nn)$ 
where $\nn\descendant n$ and $\cxy\neq\cbb$.
\end{myax}

\begin{myax}\label{ax_assert}
Assertion commands can be removed:
$\cxy(n,V_Y) \eqext \emptyseq$ where $\cxy=\cbb$ or $\cxy=\cdd$.
\end{myax}

@@

A sequence that transforms $\FS$ into $\FS_A$ must contain at least one command on each node
that changes between the two states of the filesystem. Also, it is easy to see that this is also
sufficient, as there are commands available for any input and output value. 
Specifically, if at node $n$, $\FS(n)=X'$ and $\FS_A(n)=Y'\neq X'$, then we add the command $\cxy(n,Y')$ to $A$
where $X'\in\setvx{X}$ and $Y'\in\setvx{Y}$.

@@

Let $A$ be a sequence that satisfies the conditions;
we therefore know that it contains a command $\cxy(n,V_Y)$ on $n$
and another command $\czw(\nn,V_W)$ on $\nn$.
We use an indirect proof and assume that there are no commands on $\parent(n)$ in $A$.
Next, we create a new sequence $A'\equiv A$ in which $\cxy$ and $\czw$ are next to each other.
If they are already next to each other in $A$, there is nothing to do.
Otherwise, consider the command next to $\cxy(n,V_Y)$ in the direction where $\czw(\nn,V_W)$ is.
Let this command be $\cqr(m,V_R)$.
If $\czw$ is to the right, then $A$ looks like the following:
\[ A = \cdots\cc\cxy(n,V_Y)\cc\cqr(m,V_R)\cc\cdots\cc\czw(\nn,V_W)\cc\cdots \]
If $n\unrel m$, swap $\cxy$ and $\cqr$. Based on \cref{ax_separate_commute} we know that the new
sequence is equivalent to $A$.
Otherwise, we know $m\neq\parent(n)$ as there are no commands on $\parent(n)$, and so
from \cref{ax_distantrel_breaks} we get $A\equiv\cbrk$ which contradicts our assumptions.
(Note that $A$ cannot contains assertion commands.)
By repeating this step we can therefore convert $A$ into $A'$ where $\cxy$ and $\czw$ are neighboring commands.
However, then \cref{ax_distantrel_breaks} applies to the sequence and therefore $A\equiv A'\equiv\cbrk$ which
is again a contradiction.
\end{proof}

@@

As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.
An important property of simple sequences is that
all of their valid permutations are equivalent,
and all equivalent sequences are mere permutations:

@@

\begin{mydef}{Independent commands, sequences and sets.}
For two commands $\cxy$ and $\czw$ we write $\cxy\indep\czw$ if $\cxy\cc \czw\equiv\czw\cc \cxy\nequiv\cbrk$.
For two sequences or sets of commands $A$ and $B$ we write $A\indep B$ if
$\forall a\in A \mbox{~and~} \forall b\in B: a\indep b$.
\end{mydef}

@@

\begin{mylem}\label{unrelated_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are unrelated:
\[ \cxy(n,V_Y)\indep\czw(m,V_W) \Leftrightarrow n\unrel m \]
if $\cxy(n,V_Y)\neq\czw(m,V_W)$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
We proceed in two steps; first, we prove that
\[ \cxy(n,V_Y)\nindep\czw(m,V_W) \wedge n\unrel m \] would lead to contradiction.
By definition, if the two commands are not independent, then at least one of the following holds:
\begin{itemize}
\item $\czw(m,V_W)\cc \cqr(o,V_R) \equiv\cbrk$
\item $\cqr(o,V_R)\cc \czw(m,V_W) \equiv\cbrk$
\item $\czw(m,V_W)\cc \cqr(o,V_R) \nequiv \cqr(o,V_R)\cc \czw(m,V_W)$
\end{itemize}
However, from \cref{ax_separate_commute,ax_separate_nobreaks} we also know that
if $n\unrel m$ then none of the above can be true.

Next, we prove that
\[ \cxy(n,V_Y)\indep\czw(m,V_W) \wedge n\nunrel m \] would lead to contradiction.
As $n\nunrel m$
we know that $n=m$, $n\descendant m$ or $m\descendant n$.

If $n=m$, then from \cref{ax_same_breaks} we know $Y=Z$ and $W=X$
as otherwise the commands, in one order or the other, would break all filesystems
and they could not be independent.
The commands are not assertion commands, and so from \cref{ax_same_singlec}
we know that $\cxy(n,V_Y)\cc\czw(n,V_W)\equiv\czw(n,V_W)\cc\cxy(n,V_Y)$ can only happen if $Y=W$ and $V_Y=V_W$.
Therefore $X=Y=Z=W$ which is only possible if both commands are $\cff$ and have the same
output value. This, however, contradicts our assumption that they are different.

If $n\descendant m$ or $m\descendant n$, then
from \cref{ax_distantrel_breaks} we know that if they are not directly related,
then $\cxy(n,V_Y)\cc\czw(m,V_W)$ breaks all filesystems, and they cannot be independent.
From the construction and destruction pairs and 
\cref{ax_directchild_breaks,ax_directparent_breaks} we also see that
even if they are directly related, either
$\cxy(n,V_Y)\cc\czw(m,V_W)$ or $\czw(m,V_W)\cc\cxy(n,V_Y)$ 
breaks all filesystems, so they again cannot be independent.
\end{proof}

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\begin{align*}
\forall\FS: &\cxy(n,C_Y)\indep \czw(m,C_W) \\
&\quad\wedge \cxy(n,C_Y)\FS\neq\fsbroken \\
&\quad\wedge \czw(m,C_W)\FS\neq\fsbroken \\
&\Rightarrow \cxy(n,C_Y)\cc \czw(m,C_W)\FS\neq\fsbroken.
\end{align*}
\end{myax}
\begin{proof}
We name this proposition an axiom for the same reason as above,
but it is easy to provide an indirect proof based on our filesystem model.
Assume that $\cxy(n,C_Y)\cc \czw(m,C_W)\FS=\fsbroken$.
We know $\cxy(m,C_Y)\FS\neq\fsbroken$ so it must be applying 
$\czw(m,C_W)$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $\FS(m)\in\setvx{W}$ and so
$(\cxy(m,C_Y)\FS)(m)\in\setvx{W}$ as based on \cref{unrelated_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxy(m,C_Y)$.
\end{proof}

@@

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on node $n$ in both $A$ ($\cxy(n, V_Y)\in A$) and $B$ ($\czw(n, V_W)\in B$)
then their input types must match ($X=Z$).
\end{mylem}
\begin{proof}
This is easily shown using an indirect proof: if $X\neq Z$, then there is no filesystem that
either $\cxy(n, V_Y)$ or $\czw(n, V_W)$ would not break, 
and consequently $A$ and $B$ cannot work on the same filesystem.
\end{proof}

>>>>

\newcommand{\setv}{\mathcal{V}} % Set of values
\newcommand{\setvx}[1]{\mathcal{V}_{#1}} % Set of values with custom index
\newcommand{\setf}{\setvx{F}} % Set of file values
\newcommand{\setd}{\setvx{D}} % Set of directory values
\newcommand{\setb}{\setvx{\empt}} % Set of empty values
\newcommand{\setn}{\mathcal{N}} % Set of nodes / paths
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\topnode}{\mathtt{None}} % return value of parent()
\newcommand{\fsbroken}{\mathtt{Broken}} % broken FS value
\newcommand{\FS}{\Phi} % a filesystem
\newcommand{\GS}{\Xi} % another filesystem
\newcommand{\nn}{n^*} % parent node

\newcommand{\cbrk}{\mathtt{break}} % 'break' command
\newcommand{\ccharb}{\empt}
\newcommand{\ccharf}{\mathtt{F}}
\newcommand{\cchard}{\mathtt{D}}
\newcommand{\vald}{d} % Directory value
\newcommand{\valdi}{d_1}
\newcommand{\valdii}{d_2}
\newcommand{\valf}{f} % File value
\newcommand{\valvx}{v_X} % Value of type X
\newcommand{\valvy}{v_Y} % Value of type Y
\newcommand{\valvw}{v_W}
\newcommand{\valvr}{v_R}
\newcommand{\typeset}{\{\cchard,\ccharf,\ccharb\}} % Set of types in commands
\newcommand{\caa}[2]{\langle{#1,#2}\rangle}
\newcommand{\caaa}[3]{\langle{#1,#2,#3}\rangle}
\newcommand{\caaaa}[4]{\langle{#1,#2,#3,#4}\rangle}
\newcommand{\cbb}{\caa{\ccharb}{\ccharb}}
\newcommand{\cbba}[1]{\caaa{\ccharb}{\ccharb}{#1}}
\newcommand{\cbf}{\caa{\ccharb}{\ccharf}}
\newcommand{\cbfaa}[2]{\caaaa{\ccharb}{\ccharf}{#1}{#2}}
\newcommand{\cbd}{\caa{\ccharb}{\cchard}}
\newcommand{\cbdaa}[2]{\caaaa{\ccharb}{\cchard}{#1}{#2}}
\newcommand{\cfb}{\caa{\ccharf}{\ccharb}}
\newcommand{\cfba}[1]{\caaa{\ccharf}{\ccharb}{#1}}
\newcommand{\cff}{\caa{\ccharf}{\ccharf}}
\newcommand{\cfd}{\caa{\ccharf}{\cchard}}
\newcommand{\cfdaa}[2]{\caaaa{\ccharf}{\cchard}{#1}{#2}}
\newcommand{\cdb}{\caa{\cchard}{\ccharb}}
\newcommand{\cdba}[1]{\caaa{\cchard}{\ccharb}{#1}}
\newcommand{\cdf}{\caa{\cchard}{\ccharf}}
\newcommand{\cdfaa}[2]{\caaaa{\cchard}{\ccharf}{#1}{#2}}
\newcommand{\cdd}{\caa{\cchard}{\cchard}}
\newcommand{\cdda}[1]{\caaa{\cchard}{\cchard}{#1}}
\newcommand{\cxy}{\caa{X}{Y}}
\newcommand{\cxyaa}[2]{\caaaa{X}{Y}{#1}{#2}}
\newcommand{\cxynv}{\caaaa{X}{Y}{n}{\valvy}}
\newcommand{\cxynnv}{\caaaa{X}{Y}{\nn}{\valvy}}
\newcommand{\cxw}{\caa{X}{W}}
\newcommand{\cxwnv}{\caaaa{X}{W}{n}{\valvw}}
\newcommand{\czw}{\caa{Z}{W}}
\newcommand{\czwnv}{\caaaa{Z}{W}{n}{\valvw}}
\newcommand{\czwnnv}{\caaaa{Z}{W}{\nn}{\valvw}}
\newcommand{\czwmv}{\caaaa{Z}{W}{m}{\valvw}}
\newcommand{\cqr}{\caa{Q}{R}}
\newcommand{\cqrmv}{\caaaa{Q}{R}{m}{\valvr}}
\newcommand{\cqrov}{\caaaa{Q}{R}{o}{\valvr}}

\newcommand{\cc}{\circ} % Command / sequence concatenation
\newcommand{\descendant}{\prec}
\newcommand{\descendantEq}{\preceq}
\newcommand{\ancestor}{\succ}
\newcommand{\ancestorEq}{\succeq}

@@

In this paper, we write $\valf$ for an arbitrary element of $\setf$, and $\vald$ for an arbitrary element
of $\setd$. $TODO$ is a value from $\setv$ and $n$ and $m$ are nodes in $\setn$.

@@

% TODO: tuple / quadruple
We therefore propose to have a command for each pair of types.
For want of a better system, we will name our commands by concatenating
two of $\cchard$, $\ccharf$ and $\ccharb$, 
where the left sign denotes the type of value
required in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example, $mkdir(n,D)$ is $\cbdaa{n}{D}$, and $rmdir(n)$ is $\cdba{n}$.

@@

We note that a command or a sequence of commands is applied to a filsystem
by prefixing the command or sequence to it, for example: $\cbrk\FS$, $\cbdaa{n}{\vald}\FS$, 
or $S\FS$ if $S$ is a sequence of commands.

@@

The exact behavior of our commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
All other commands have the generic form $\cxynv$ where
$X$ and $Y$ are types in $\typeset$,
$n\in\setn$ and $\valvy$ is a value of the appropriate type from $\setvx{Y}$.
At times we omit $\valvy$ if there is only one suitable value.
\[ \cxynv\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[V/n] \mbox{~violates the tree property}\\
   \FS[V/n] &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

In general, when we describe multiple or unknown commands, we may substitute one or both
of their types with variables, as we did with $\cxynv$. In this context,
if we write that $\cxy=\czw$, we mean that the input and output types % TODO needed?
of the commands are the same ($X=Z$ and $Y=W$), while a full equiality 
($\cxynv=\czwmv$)
implies
that their output value is also the same.

@@

The commands we aim to investigate do not occur in isolation:
for example, when comparing filesystems to determine the commands needed to transform one into the other,
we receive a set of commands with uncertain ordering.
Applying commands to a filesystem, however, happens in time,
so for this we need a sequence of commands.
We use $\cc$ to concatenate commands to form a sequence, or concatenate sequences to form a longer sequence:
$S=\cbdaa{n}{\vald}\cc\cbfaa{m}{\valf}$, or $S'=S\cc\cbba{o}$.
We write $\emptyseq$ for the empty sequence.

@@

Pairs of commands in general have the form
\[ \cxynv\cc  \czwmv \]
where $X,Y,Z,W\in\typeset$, $n,m\in\setn$, 
and values are of the appropriate type: 
$\valvy\in\setvx{Y}$ and $\valvw\in\setvx{W}$.

\begin{myax}\label{ax_separate_commute}
Commands on unrelated nodes commute:
$\cxynv\cc\czwmv \equiv \czwmv\cc\cxynv$ where $n\unrel m$.
\end{myax}

\begin{myax}\label{ax_separate_nobreaks}
Commands on unrelated nodes do not break every filesystem:
$\cxynv\cc\czwmv \nequiv \cbrk$ where $n\unrel m$.
\end{myax}

\begin{myax}\label{ax_same_breaks}
Commands on the same node break every filesystem if their types are incompatible:
$\cxynv\cc\czwnv \equiv \cbrk$ where $Y\ne Z$.
\end{myax}

\begin{myax}\label{ax_same_emptyseq}
Commands on the same node simplify:
$\cxynv\cc\czwnv \eqext \emptyseq$ where $Y=Z$ and $X=W=\ccharb$ 
or $X=W=\cchard$.
\end{myax}

\begin{myax}\label{ax_same_singlec}
Commands on the same node simplify:
$\cxynv\cc \czwnv \equiv \cxwnv$ where $Y=Z$ and $\cxw\neq\cbb$ and $\cxw\neq\cdd$.
\end{myax}

\begin{myax}\label{ax_distantrel_breaks}
Commands on distant relatives break all filesystems:
$\cxynnv\cc\czwnv \equiv \cbrk$
and $\czwnv\cc\cxynnv \equiv\cbrk$
where $\nn\descendant n$ and $\nn\neq\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$.
\end{myax}

\begin{mydef}{Construction pair.}
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parent(n)$ 
and the commands are one
of the following:
   \begin{gather*}
   % bd>bf
            \cbdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
   % bd>bd
            \cbdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii} \\
   % fd>bf
            \cfdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
   % fd>bd
            \cfdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii}
   \end{gather*}
\end{mydef}

\begin{myax}\label{ax_directchild_breaks}
All other commands on a child break every filesystem:
$\cxynnv\cc\czwnv \equiv \cbrk$ where $\nn=\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.
\end{myax}

\begin{mydef}{Destruction pair.}
A pair of commands on nodes $n$ and $\nn$ is a destruction pair if $\parent(n)=\nn$ and the commands are one
of the following:
   \begin{gather*}
   % fb<db ??
            \cfba{n}\cc  \cdba{\nn} \\
   % fb<df ??
            \cfba{n}\cc  \cdfaa{\nn}{\valf} \\
   % db<db ??
            \cdba{n}\cc  \cdba{\nn} \\
   % db<df ??
            \cdba{n}\cc  \cdfaa{\nn}{\valf}
   \end{gather*}
\end{mydef}

\begin{myax}\label{ax_directparent_breaks}
All other commands on a parent break every filsystem:
$\cxynv\cc\czwnnv \equiv \cbrk$ where $\parent(n)=\nn$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.
\end{myax}

\begin{mydef}{Assertion command.}
A command is an assertion command if
for every filesystem it either breaks it or leaves it in the same state.
In other words, assertion commands have the same input and output type,
and there must also be only one possible value for their output.
Accordingly, in our current algebra, $\cbba{n}$ and $\cdda{n}$ are the only two assertion commands,
and we do not note their output value as it is determined.
\end{mydef}

\begin{myax}\label{ax_child_assert}
An assertion command can be added on a descendant node:
$\cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n}$ 
where $\nn\descendant n$ and $\cxy\neq\cdd$.
\end{myax}

\begin{myax}\label{ax_parent_assert}
An assertion command can be added on an ancestor node:
$\cdda{\nn}\cc\cxynv \equiv \cxynv \equiv \cxynv\cc\cdda{\nn}$ 
where $\nn\descendant n$ and $\cxy\neq\cbb$.
\end{myax}

\begin{myax}\label{ax_assert}
Assertion commands can be removed:
$\cxynv \eqext \emptyseq$ where $\cxy=\cbb$ or $\cxy=\cdd$.
\end{myax}

@@

A sequence that transforms $\FS$ into $\FS_A$ must contain at least one command on each node
that changes between the two states of the filesystem. Also, it is easy to see that this is also
sufficient, as there are commands available for any input and output value. 
Specifically, if at node $n$, $\FS(n)=X'$ and $\FS_A(n)=Y'\neq X'$, then we add the command $\cxyaa{n}{Y'}$ to $A$
where $X'\in\setvx{X}$ and $Y'\in\setvx{Y}$.

@@

Let $A$ be a sequence that satisfies the conditions;
we therefore know that it contains a command $\cxynv$ on $n$
and another command $\czwnnv$ on $\nn$.
We use an indirect proof and assume that there are no commands on $\parent(n)$ in $A$.
Next, we create a new sequence $A'\equiv A$ in which $\cxy$ and $\czw$ are next to each other.
If they are already next to each other in $A$, there is nothing to do.
Otherwise, consider the command next to $\cxynv$ in the direction where $\czwnnv$ is.
Let this command be $\cqrmv$.
If $\czw$ is to the right, then $A$ looks like the following:
\[ A = \cdots\cc\cxynv\cc\cqrmv\cc\cdots\cc\czwnnv\cc\cdots \]
If $n\unrel m$, swap $\cxynv$ and $\cqrmv$. Based on \cref{ax_separate_commute} we know that the new
sequence is equivalent to $A$.
Otherwise, we know $m\neq\parent(n)$ as there are no commands on $\parent(n)$, and so
from \cref{ax_distantrel_breaks} we get $A\equiv\cbrk$ which contradicts our assumptions.
(Note that $A$ cannot contains assertion commands.)
By repeating this step we can therefore convert $A$ into $A'$ where $\cxynv$ and $\czwnnv$ are neighboring commands.
However, then \cref{ax_distantrel_breaks} applies to the sequence and therefore $A\equiv A'\equiv\cbrk$ which
is again a contradiction.
\end{proof}

@@

As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.
An important property of simple sequences is that
all of their valid orderings are equivalent,
and all equivalent sequences are mere reorderings:

@@

\begin{mydef}{Independent commands, sequences and sets.}
For two commands $\cxynv$ and $\czwmv$ we write $\cxynv\indep\czwmv$ if $\cxynv\cc\czwmv \equiv \czwnv\cc\cxynv \nequiv \cbrk$.
For two sequences or sets of commands $A$ and $B$ we write $A\indep B$ if
$\forall a\in A \mbox{~and~} \forall b\in B: a\indep b$.
\end{mydef}

@@

\begin{mylem}\label{unrelated_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are unrelated:
\[ \cxynv\indep\czwmv \Leftrightarrow n\unrel m \]
if $\cxynv\neq\czwmv$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
We proceed in two steps; first, we prove that
\[ \cxynv\nindep\czwmv \wedge n\unrel m \] would lead to contradiction.
By definition, if the two commands are not independent, then at least one of the following holds:
\begin{itemize}
\item $\czwmv\cc \cqrov \equiv\cbrk$
\item $\cqrov\cc \czwmv \equiv\cbrk$
\item $\czwmv\cc \cqrov \nequiv \cqrov\cc \czwmv$
\end{itemize}
However, from \cref{ax_separate_commute,ax_separate_nobreaks} we also know that
if $n\unrel m$ then none of the above can be true.

Next, we prove that
\[ \cxynv\indep\czwmv \wedge n\nunrel m \] would lead to contradiction.
As $n\nunrel m$
we know that $n=m$, $n\descendant m$ or $m\descendant n$.

If $n=m$, then from \cref{ax_same_breaks} we know $Y=Z$ and $W=X$
as otherwise the commands, in one order or the other, would break all filesystems
and they could not be independent.
The commands are not assertion commands, and so from \cref{ax_same_singlec}
we know that $\cxynv\cc\czwnv\equiv\czwnv\cc\cxynv$ can only happen if $Y=W$ and $\valvy=\valvw$.
Therefore $X=Y=Z=W$ which is only possible if both commands are $\cff$ and have the same
output value. This, however, contradicts our assumption that they are different.

If $n\descendant m$ or $m\descendant n$, then
from \cref{ax_distantrel_breaks} we know that if they are not directly related,
then $\cxynv\cc\czwmv$ breaks all filesystems, and they cannot be independent.
From the construction and destruction pairs and 
\cref{ax_directchild_breaks,ax_directparent_breaks} we also see that
even if they are directly related, either
$\cxynv\cc\czwmv$ or $\czwmv\cc\cxynv$ 
breaks all filesystems, so they again cannot be independent.
\end{proof}

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\begin{align*}
\forall\FS: &\cxynv\indep \czwmv \\
&\quad\wedge \cxynv\FS\neq\fsbroken \\
&\quad\wedge \czwmv\FS\neq\fsbroken \\
&\Rightarrow \cxynv\cc \czwmv\FS\neq\fsbroken.
\end{align*}
\end{myax}
\begin{proof}
We name this proposition an axiom for the same reason as above,
but it is easy to provide an indirect proof based on our filesystem model.
Assume that $\cxynv\cc \czwmv\FS=\fsbroken$.
We know $\cxynv\FS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $\FS(m)\in\setvx{W}$ and so
$(\cxynv\FS)(m)\in\setvx{W}$ as based on \cref{unrelated_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

@@

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on node $n$ in both $A$ ($\cxynv\in A$) and $B$ ($\czwnv\in B$)
then their input types must match ($X=Z$).
\end{mylem}
\begin{proof}
This is easily shown using an indirect proof: if $X\neq Z$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $A$ and $B$ cannot work on the same filesystem.
\end{proof}
